Einstein's gravity from a polynomial affine model

### 2016-05-10 AffGR-PRD.tex
Reviewed until propagators section

Until now, the most general diffeomorphism-invariant and power-counting renormalizable (gravitational) theory for an affine connection has been built, and we have showed that in a certain sector it is equivalent to General Relativity. In what follows, we show an attempt of including scalar matter into the model. The theory does not require a spacetime metric, but instead we should consider the most general $\binom{2}{0}$-tensor density, $\mathfrak{g}^{\mu\nu}$,  built with the available fields, and use it to build Lagrangian densities for the matter content. Following our precept, the matter content---scalar matter---should couple to $\mathfrak{g}^{\mu\nu}$, given by
\begin{dmath}
  \mathfrak{g}^{\mu\nu} = \alpha \, \nabla_\lambda T^{\mu,\nu\lambda} + \beta \, A_\lambda T^{\mu,\nu\lambda} + \gamma \, \epsilon_{\lambda\kappa\rho\sigma} T^{\mu, \lambda\kappa} T^{\nu, \rho\sigma},
  \label{geng}
\end{dmath}
with $\alpha$, $\beta$ and $\gamma$ arbitrary coefficients.

We consider the action provided by the ``kinetic term''
\begin{dmath}
  \label{ScalarAction}
  S_\phi = -  \int \dn{4}{x} \Big( \alpha \, \nabla_\lambda T^{\mu,\nu\lambda}  + \beta \, A_\lambda T^{\mu,\nu\lambda} + \gamma \, \epsilon_{\lambda\kappa\rho\sigma} T^{\mu, \lambda\kappa} T^{\nu, \rho\sigma} \Big) \partial_\mu\phi\partial_\nu\phi,
\end{dmath}
which makes a nontrivial contribution to the field equations once we restrict to the sector of interest, where the torsion is set to zero and the connection is compatible with a metric.

@@

Therefore, in notation of Eqs.~\eqref{GammaExp} we have that $A = T = 0$, and the contribution due to the torsion (and non-metricity) enters as a perturbative effect. Thus, most terms in the action~\eqref{4dfull} do not contribute to the propagator analysis. Additionally, the first order perturbation of the curvature tensor is

@@

Thus, the perturbations can be split into their longitudinal and transverse components as
\begin{dmath}
  a_\mu = {a_\perp}_{\mu} + {\nabla}_{\mu}{a_{\mathtt{L}}},
\end{dmath}
\begin{dmath}
  {t}^{\lambda \mu \nu} =
  {t_\perp}^{\lambda \mu \nu}
  + 2  {\nabla}^{\lambda} {t_{\mathtt{a}}^{\mu \nu}}
  - 2  {\nabla}^{[\mu}{{t_{\mathtt{a}}}^{\nu] \lambda}} 
  + 2  {\nabla}^{\mu}{{t_{\mathtt{s}}}^{\nu \lambda}}
  + 2  {\nabla}^{\lambda [\mu}{{t_{\mathtt{v}}}^{\nu]}}
\end{dmath}
and
\begin{dmath}
  {\gamma}^{\lambda}\,_{\mu \nu}  =
  {\gamma_\perp}^{\lambda}{}_{\mu \nu}
  + {\nabla}^{\lambda}{{\gamma_{\mathtt{s}}}_{\mu \nu}}
  + 2  {\nabla}_{(\mu}{{\gamma}^{\lambda}\,_{\nu)}}
  + 2  {\nabla}^{\lambda}\nabla_{(\mu}{{\gamma_{\mathtt{v}}}_{\nu)}}
  + 2  \nab{(\mu} \nab{\nu)} {{\gamma_{\mathtt{v}'}}^{\lambda}} 
  + 2  {\nabla}^{\lambda}\nab{(\mu} \nab{\nu)}{\gamma_{\mathtt{L}}},
\end{dmath}
where all the indices are transverse except those of the derivative $\nabla_\mu$, the subindex $\perp$ indicates that it corresponds to the fully transverse projection of the tensor in all of its indices, the subindex ${\mathtt{L}}$ indicates that this is the longitudinal projection of the tensor, while the sub indices $\mathtt{s}$, $\mathtt{a}$ and $\mathtt{v}$ specify that the tensor is symmetric, antisymmetric or a vector representation.
%%In particular, $t_\perp^{\lambda}{}_\kappa$ is traceless. \hl{check the last sentence. Is this ${t_s}^{\alpha}{}_{\alpha}$?... If true, this condition will kill the scalars contribution in the next equations}

>>>>

Until now, the most general diffeomorphism-invariant and power-counting renormalizable (gravitational) theory for an affine connection has been built, and we have showed that in a certain sector it is a generalization of General Relativity. In what follows, we show an attempt of including scalar matter into the model. The theory does not require a spacetime metric, but instead we should consider the most general, symmetric $\binom{2}{0}$-tensor density, $\mathfrak{g}^{\mu\nu}$,  built with the available fields (see Appendix~\ref{sec:im}), and use it to build Lagrangian densities for the matter content. Following our precept, the matter content---scalar matter---should couple to $\mathfrak{g}^{\mu\nu}$, given by
\begin{dmath}
  \mathfrak{g}^{\mu\nu} = \alpha \, \nabla_\lambda T^{\mu,\nu\lambda} + \beta \, A_\lambda T^{\mu,\nu\lambda} + \gamma \, \epsilon_{\lambda\kappa\rho\sigma} T^{\mu, \lambda\kappa} T^{\nu, \rho\sigma},
  \label{geng}
\end{dmath}
with $\alpha$, $\beta$ and $\gamma$ arbitrary coefficients.

We consider the action provided by the ``kinetic term''
\begin{dmath}
  \label{ScalarAction}
  S_\phi = -  \int \dn{4}{x} \Big( \alpha \, \nabla_\lambda T^{\mu,\nu\lambda}  + \beta \, A_\lambda T^{\mu,\nu\lambda} + \gamma \, \epsilon_{\lambda\kappa\rho\sigma} T^{\mu, \lambda\kappa} T^{\nu, \rho\sigma} \Big) \partial_\mu\phi\partial_\nu\phi,
\end{dmath}
which makes a nontrivial contribution to the field equations once we restrict to the sector of interest.
%%, where the torsion is set to zero and the connection is compatible with a metric.

@@

Therefore, in notation of Eqs.~\eqref{GammaExp} we have that $A = T = 0$, and the contribution due to the torsion and non-metricity enters as a perturbative effect. Thus, most terms in the action~\eqref{4dfull} do not contribute to the propagator analysis. Additionally, the first order perturbation of the curvature tensor is

@@

Thus, the perturbations can be split into their longitudinal and transverse components as
\begin{dmath}
  a_\mu = {a_\perp}_{\mu} + {\nabla}_{\mu}{a_{\mathtt{L}}},
\end{dmath}
\begin{dmath}
  {t}^{\lambda \mu \nu} =
  {t_\perp}^{\lambda \mu \nu}
  + 2  {\nabla}^{\lambda} {t_{\mathtt{a}}^{\mu \nu}}
  - 2  {\nabla}^{[\mu}{{t_{\mathtt{a}}}^{\nu] \lambda}} 
  + 2  {\nabla}^{\mu}{{t_{\mathtt{s}}}^{\nu \lambda}}
  + 2  {\nabla}^{\lambda [\mu}{{t_{\mathtt{v}}}^{\nu]}}
\end{dmath}
and
\begin{dmath}
  {\gamma}^{\lambda}\,_{\mu \nu}  =
  {\gamma_\perp}^{\lambda}{}_{\mu \nu}
  + {\nabla}^{\lambda}{{\gamma_{\mathtt{s}}}_{\mu \nu}}
  + 2  {\nabla}_{(\mu}{{\gamma}^{\lambda}\,_{\nu)}}
  + 2  {\nabla}^{\lambda}\nabla_{(\mu}{{\gamma_{\mathtt{v}}}_{\nu)}}
  + 2  \nab{(\mu} \nab{\nu)} {{\gamma_{\mathtt{v}'}}^{\lambda}} 
  + 2  {\nabla}^{\lambda}\nab{(\mu} \nab{\nu)}{\gamma_{\mathtt{L}}},
\end{dmath}
where all the indices are transverse except those of the derivative $\nabla_\mu$, the subindex $\perp$ indicates that it corresponds to the fully transverse projection of the tensor in all of its indices, the subindex ${\mathtt{L}}$ indicates that this is the longitudinal projection of the tensor, while the subindices $\mathtt{s}$, $\mathtt{a}$ and $\mathtt{v}$ specify whether the tensor is symmetric, antisymmetric or a vector representation.
%%In particular, $t_\perp^{\lambda}{}_\kappa$ is traceless. \hl{check the last sentence. Is this ${t_s}^{\alpha}{}_{\alpha}$?... If true, this condition will kill the scalars contribution in the next equations}
